API design for machine learning software: experiences from the scikit-learn project

### 2013-06-25 paper.tex
add astroML reference; fix typos; update user statistics

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g., machine learning,
natural language processing, or neuroscience). The project also benefits
from many occasional contributors proposing small code or documentation fixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Complete documentation and examples are required for all features,
and major modifications must pass code review by at least two
developers not involved in the implementation of the proposed change.

Over the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gaged from various indicators such as hundreds
of citations in scientific publications, successes in various machine learning
challenges (e.g., Kaggle), or significant statistics derived from our
repositories and mailing list.  At the time of writing, the project is watched
by 1309 people and forked 667 times on GitHub; the mailing list receives more
than 300 mails per month; version control logs
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 unique contributors to the codebase and the online documentation
website receives 37k unique visitors and 295k page views per month.

@@

While part of the scikit-learn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The simplicity, power, and extensibility of the scikit-learn API is evidenced
by the large and growing user-base, the use of scikit-learn to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the scikit-learn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}, a package providing
machine learning tools for astronomers and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of 
the recently-proposed sparse multiclass
algorithm \citep{mblondel-mlj2013}, released as
part of the 
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the scikit-learn conventions. To maximize ease of use, we
encourage more researchers to follow the scikit-learn conventions when
releasing their software in the future.

>>>>

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g., machine learning,
natural language processing, neuroscience, astrophysics, etc.). The project also benefits
from many occasional contributors proposing small code or documentation fixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Complete documentation and examples are required for all features,
and major modifications must pass code review by at least two
developers not involved in the implementation of the proposed change.

Over the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gauged from various indicators such as the hundreds
of citations in scientific publications, successes in various machine learning
challenges (e.g., Kaggle), or significant statistics derived from our
repositories and mailing list.  At the time of writing, the project is watched
by 1365 people and forked 693 times on GitHub; the mailing list receives more
than 300 mails per month; version control logs
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 unique contributors to the codebase and the online documentation
website receives 37,000 unique visitors and 295,000 page views per month.

@@

While part of the scikit-learn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The simplicity, power, and extensibility of the scikit-learn API is evidenced
by the large and growing user-base, the use of scikit-learn to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the scikit-learn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of 
the recently-proposed sparse multiclass
algorithm \citep{mblondel-mlj2013}, released as
part of the 
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the scikit-learn conventions. To maximize ease of use, we
encourage more researchers to follow the scikit-learn conventions when
releasing their software in the future.
